\textls[-0]{\thetitle}

### 2023-03-18 metatheory2.tex
Add details about Grothendieck Universe
edit in appendix

In this section, we lay out the complete version of the syntax of \TT but in the style of \citeauthor{debruijn1964} indices and \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., in the form of \sub{T}{\gm}) are part
of the syntax rather than meta-operations. What's more, we also make universe level explicit for completeness.

Other parts of the syntax remained the same as the main text.

We also work in an intrinsic-typed setting, so there is no un-typed preterms appearing anywhere. We introduce judgemental equality ($\equiv$)
for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
contexts and explicits substitutions.

This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT.

After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.

Again, following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}, we consider our syntax as the result of taking quotient by all of these judgemental equalities. This means 
(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.

@@

% We first need the mathmatical setup to talk about universe level, following \citet{sterling2019algebraic}:

% \newcommand{\Set}[1]{\ensuremath{\texttt{Set}_{#1}}}

% {\textbf{Set-theoretic Universe Assumption.}} As a simplifying move, we assume an infinite hierarchy of Grothendieck universes $\Set{i}$ for $i \in \mathbb{N}$ in our meta-logic.

% For people familiar with Agda, we can consider $\Set{i}$ as the one in Agda. In other words, each $\Set{i}$ is (1). closed under dependent function type and dependent tuple type, (2). cumulative as $\Set{i} \in \Set{i+1}$ and $\Set{i} \subseteq \Set{i+1}$

>>>>

This appendix supplements \cref{sec:metatheory2} with the details of the syntax
and the proofs of the main results.

Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
use \citeauthor{debruijn1964} indices and \emph{explicit
substitutions}~\cite{abadi1989subst,substcalculus}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
applies~$\gm$ to the type~$T$) are part of the syntax rather than
meta-operations. 
We work in an intrinsically typed setting: terms are well typed by construction.
This style of syntax formulation follows a recent
trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
known as the ``algebraic presentation'' of MLTT.
%We introduce judgmental equality ($\equiv$)
%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
%contexts, explicits substitutions, and etc. 
%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
Moreover, universe levels are explicit in this appendix.

Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
In particular,
%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
the syntax is still considered as being quotiented by the judgmental equalities.
Quotienting facilitates coercion along equalities.
Furthermore, the semantic model we develop in this appendix respects these
judgmental equalities by construction.
%This means 
%(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.

%After the syntax, we detail the models and proofs for syntactic translation (that translate away linkages),  consistency and canonicity.

@@
